---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
act <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
1. Make a histogram of the total number of steps taken each day
```{r}
act_sum_date <- aggregate(steps ~ date, data=act, sum)
hist(act_sum_date$steps, main="Total Number of Steps Taken per Day", xlab="Number of Steps")
```

2. Calculate and report the mean and median total number of steps taken
per day
```{r}
act_mean_date <- aggregate(steps ~ date, data=act, mean)
colnames(act_mean_date) <- c("date", "average_steps")
act_mean_date
```

```{r}
act_median_date <- aggregate(steps ~ date, data=act, median)
colnames(act_median_date) <- c("date", "median_steps")
act_median_date
```

## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
and the average number of steps taken, averaged across all days (y-axis)
```{r}
act_mean_interval <- aggregate(steps ~ interval, data=act, mean)
colnames(act_mean_interval) <- c("interval", "average_steps")

plot(act_mean_interval, type="l", main="Daily Average Steps per 5-minutes Interval")
```

2. Which 5-minute interval, on average across all the days in the dataset,
contains the maximum number of steps?
```{r}
subset(act_mean_interval, average_steps==max(act_mean_interval$average_steps))

```


## Imputing missing values

1. Calculate and report the total number of missing values in the dataset.
```{r}
nrow(act[is.na(act$steps),])
```

2. Devise a strategy for filling in all of the missing values in the dataset.  
    Missing values will be replaced with the average steps of the interval.

3. Create a new dataset that is equal to the original dataset but with the
missing data filled in.
```{r}
act_na <- act[is.na(act$steps),]
act_na_average <- merge(act_na, act_mean_interval)
act_na_average$steps <- act_na_average$average_steps
act_filled <- rbind(act_na_average[, c("steps", "date", "interval")], act[!is.na(act$steps), ])
```

4. Make a histogram of the total number of steps taken each day and Calculate
and report the mean and median total number of steps taken per day.
```{r}
act_filled_sum_date <- aggregate(steps ~ date, data=act_filled, sum)
hist(act_filled_sum_date$steps, main="Total Number of Steps Taken per Day with Missing value Replaced", xlab="Number of Steps")
```

```{r}
act_filled_mean_date <- aggregate(steps ~ date, data=act_filled, mean)
colnames(act_filled_mean_date) <- c("date", "average_steps")
act_filled_mean_date
```

```{r}
act_filled_median_date <- aggregate(steps ~ date, data=act_filled, median)
colnames(act_filled_median_date) <- c("date", "median_steps")
act_filled_median_date
```


## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels - "weekday"
and "weekend" indicating whether a given date is a weekday or weekend
day.
```{r}
act_filled_weekdays <- transform(act_filled, day=weekdays(as.Date(date)))
weekday <- data.frame(day=c("Saturday", "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), weekday=c(rep("weekend", 2), rep("weekday", 5)))
act_filled_weekdays <- merge(weekday, act_filled_weekdays)

```
2. Make a panel plot containing a time series plot (i.e. type = "l") of the
5-minute interval (x-axis) and the average number of steps taken, averaged
across all weekday days or weekend days (y-axis).
```{r}
act_filled_weekdays_mean <- aggregate(steps ~ weekday + interval, data=act_filled_weekdays, mean)
colnames(act_filled_weekdays_mean) <- c("weekday", "interval", "average_steps")
par(mfrow=c(2, 1))
plot(act_filled_weekdays_mean[act_filled_weekdays_mean$weekday=="weekend", c("interval", "average_steps")], type="l", main="Daily Average Steps per 5-minutes Interval during Weekends")
plot(act_filled_weekdays_mean[act_filled_weekdays_mean$weekday=="weekday", c("interval", "average_steps")], type="l", main="Daily Average Steps per 5-minutes Interval during Weekdays")

```

